

# Fundamentals of Linear Regression

```{r set-options, echo = FALSE, cache = FALSE}
options(width = 110)
```

> Linear regression is the foundation of most of the methods [Hayes described in his] book, so a solid understanding of the fundamentals of linear regression is essential. I assume that most readers have been exposed to linear regression in some form before discovering this book, and so some of the material will be review. Even so, I encourage everyone to read this chapter. (p. 30)

Since we're adding Bayes and the tidyverse into the mix, walking through this chapter will be double important, for us.

## Correlation and prediction

Here we load a couple necessary packages, load the data, and take a peek.

```{r, warning = F, message = F}
library(tidyverse)

glbwarm <- read_csv("data/glbwarm/glbwarm.csv")

glimpse(glbwarm)
```

If you are new to tidyverse-style syntax, possibly the oddest component is the pipe (i.e., `%>%`). I'm not going to explain the `%>%` in this project, but you might learn more about in [this brief clip](https://www.youtube.com/watch?v=9yjhxvu-pDg), starting around [minute 21:25 in this talk by Wickham](https://www.youtube.com/watch?v=K-ss_ag2k9E&t=1285s), or in [section 5.6.1 from Grolemund and Wickham’s *R for Data Science*](http://r4ds.had.co.nz/transform.html#combining-multiple-operations-with-the-pipe). Really, all of chapter 5 of *R4DS* is great for new R and new tidyverse users, and their chapter 3 is a nice introduction to plotting with ggplot2.

Here is our version of Figure 2.1. 

```{r, fig.width = 5, fig.height = 4}
glbwarm %>% 
  group_by(negemot, govact) %>% 
  count() %>% 
  
  ggplot(aes(x = negemot, y = govact)) +
  geom_point(aes(size = n),
             show.legend = F) +
  labs(x = expression(paste("NEGEMOT: Negative emotions about climate change (", italic("X"), ")")),
       y = expression(paste("GOVACT: Support for governmentaction (", italic("Y"), ")"))) +
  theme_bw()
```

There are other ways to handle the [overplotting issue, such as jittering](http://ggplot2.tidyverse.org/reference/position_jitter.html). 

```{r, fig.width = 5, fig.height = 4}
glbwarm %>% 
  ggplot(aes(x = negemot, y = govact)) +
  geom_jitter(height = .05, width = .05, 
              alpha = 1/2, size = 1/3) +
  labs(x = expression(paste("NEGEMOT: Negative emotions about climate change (", italic("X"), ")")),
       y = expression(paste("GOVACT: Support for governmentaction (", italic("Y"), ")"))) +
  theme_bw()
```

The `cor()` function is a simple way to compute a Pearson's correlation coefficient.

```{r}
cor(glbwarm$negemot, glbwarm$govact)
```

If you want more plentiful output, the `cor.test()` function returns a $t$-value, the degrees of freedom, the corresponding $p$-value and the 95% confidence intervals, in addition to the Pearson's correlation coefficient.

```{r}
cor.test(glbwarm$negemot, glbwarm$govact)
```

To get the Bayesian version, we'll open our focal statistical package, Bürkner's [brms](https://github.com/paul-buerkner/brms). [And I'll briefly note that you could also do many of these analyses with other packages, such as [blavaan](https://cran.r-project.org/web/packages/blavaan/index.html). I just prefer brms.]

```{r, message = F, warning = F}
library(brms)
```

We'll start simple and just use the default priors and settings, but with the addition of parallel sampling via `cores = 4`.

```{r model1, cache = T, message = F, warning = F}
model1 <- 
  brm(data = glbwarm, 
      family = gaussian,
      mvbind(negemot, govact) ~ 1,
      cores = 4)
```

```{r}
print(model1)
```

In regression models, we typically use predictor variables to explain variation in our criterion variables. It’s pretty much never the case that our predictors explain all the variation. The variation that’s left over is often called residual variation, residual variance, residuals, error, or even $\epsilon$. 

More formally, when we use likelihood-based estimators, such as maximum likelihood (popular with multilevel modeling and structural equation modeling) or Bayesian estimators, we express the models for our criterion variables in terms of likelihood functions. Probably the  most common likelihood function, and the one consistent with the models in Hayes’ text, is the Gaussian likelihood. With that likelihood we say our criterion variable $y_i$ is normally distributed with a mean $\mu$ and standard deviation $\sigma$. We might express that formally as

$$
y_i \sim \text{Normal} (\mu, \sigma),
$$

Where $\sim$ stands for "is distributed as" and $i$ indexes the $i^\text{th}$ row in the data. When we add predictors to the model, they are typically used to model the mean $\mu$. Thus, in the case there we have a sole predictor $x_i$ which varies across participants, we'd expand our model formula to

$$
\begin{align*}
y_i & \sim \text{Normal} (\mu_i, \sigma) \\
\mu_i & = \beta_0 + \beta_1 x_i,
\end{align*}
$$

where $\beta_0$ is the intercept and $\beta_1$ is the slope for predictor $x$, which varies across cases. In this formulation, $\sigma$ is the standard deviation after accounting for the systemic variation explained by $x_i$. That is, it's the residual variance (i.e., $\epsilon$), but in a standard-deviation metric. *Why in a standard deviation metric?*, you ask. There are technical reasons why brms expresses it as a standard deviation which I'm not going to go into, here. But just beware that whereas many frequentist software packages the express the residual variation in a variance metric, it's expressed in a standard-deviation metric in brms. Just go with it and move on. 

Within the brms framework, $\sigma$ of the Gaussian likelihood is considered a family-specific parameter. As it turns out, there are many other fine likelihood functions in addition to the Gaussian and not all of them have a $\sigma$ parameter. For example, there is no $\sigma$ for the Poisson distribution, which is popular likelihood function for count variables. Because Bürkner made the brms package capable of a variety of likelihood functions, it behooved him to give this section of the output a generic name.

When you have a regression model with multiple Gaussian criterion variables, those variables will typically have some degree of covariation. It's often termed residual covariance, particularly within the structural equation modeling paradigm. But when you have an intercept-only regression model with multiple variables, that residual covariance is just a covariance. And when you express those variation parameters in terms of standard deviations $\sigma$, their covariance is expressed as a correlation $\rho$. Since our multivariate model is of two variables, we have one $\rho$ parameter for the $\sigmas$s, `rescor(negemot,govact)`, which is our Bayesian analogue to the Pearson’s correlation.

To learn more about the multivariate syntax in brms, click [here](https://cran.r-project.org/web/packages/brms/vignettes/brms_multivariate.html) or code `vignette("brms_multivariate")`.

But to clarify the output:

* 'Estimate'  = the posterior mean, analogous to the frequentist point estimate,
* 'Est.Error' = the posterior $SD$, analogous to the frequentist standard error,
* 'l-95% CI'  = the lower-level of the percentile-based 95% Bayesian credible interval, and
* 'u-95% CI'  = the upper-level of the same.

## The simple linear regression model

Here is how one might get the simple OLS coefficients in base R with the `lm()` function.

```{r}
(model2 <- lm(data = glbwarm, govact ~ 1 + negemot))
```

For more detailed output, put the model object `model2` into the `summary()` function.

```{r}
summary(model2)
```

Here's the Bayesian model in brms.

```{r model2, cache = T, message = F, warning = F}
model2 <- 
  brm(data = glbwarm, 
      family = gaussian,
      govact ~ 1 + negemot,
      cores = 4)
```

There are several ways to get a brms model summary. A go-to is with the `print()` function.

```{r}
print(model2)
```

The `summary()` function works very much the same way. To get a more focused look, you can use the `posterior_summary()` function:

```{r}
posterior_summary(model2)
```

That also yields the log posterior, `lp__`, which you can learn more about [here](https://cran.r-project.org/web/packages/rstan/vignettes/rstan.html#the-log-posterior-function-and-gradient) or [here](https://www.jax.org/news-and-insights/jax-blog/2015/october/lp-in-stan-output). We won't focus on the `lp__` directly in this project. But it's influence will be lurking in the shadows.

But anyways, the `Q2.5` and `Q97.5`, are the lower- and upper-levels of the 95% credible intervals. The `Q` prefix stands for quantile ([see this thread](https://github.com/paul-buerkner/brms/issues/425)). In this case, these are a renamed version of the `l-95% CI` and `u-95% CI` columns from our `print()` output.

To make a quick plot of the regression line, one can use the convenient `marginal_effects()` function.

```{r, fig.width = 4.5, fig.height = 3}
marginal_effects(model2)
```

If you want to customize that output, you might nest it in `plot()`.

```{r, fig.width = 4.5, fig.height = 3}
plot(marginal_effects(model2),
     points = T,
     point_args = c(height = .05, width = .05, 
                    alpha = 1/2, size = 1/3))
```

It's also useful to be able to work with the output of a brms model directly. For our first step, we'll put our HMC draws into a data frame with the `posterior_samples()` function.

```{r}
post <- posterior_samples(model2)

head(post)
```

Next, we'll use the `fitted()` function to simulate model-implied summaries for the expected `govact` value, given particular predictor values. Our first model only has `negemot` as a predictor, and we'll ask for the expected `govact` values for `negemot` ranging from 0 to 7.

```{r}
nd <- tibble(negemot = seq(from = 0, to = 7, length.out = 30))

f <-
  fitted(model2,
         newdata = nd) %>% 
  data.frame() %>% 
  bind_cols(nd)

head(f)
```

The first two columns should look familiar to the output from `print(model2)`, above. The next two columns, `Q2.5` and `Q97.5`, are the lower- and upper-levels of the 95% credible intervals, like we got from `posterior_samples()`. We got the final column with the `bind_cols(nd)` code.

Here's our bespoke version of Figure 2.4.

```{r, fig.width = 5, fig.height = 4}
glbwarm %>% 
  group_by(negemot, govact) %>% 
  count() %>% 
  
  ggplot(aes(x = negemot)) +
  geom_point(aes(y = govact, size = n),
             show.legend = F) +
  geom_ribbon(data = f,
              aes(ymin = Q2.5,
                  ymax = Q97.5),
              fill = "grey75", alpha = 3/4) +
  geom_line(data = f,
            aes(y = Estimate)) +
  annotate("text", x = 2.2, y = 7.5, label = "Cases with positive residuals", color = "red3") +
  annotate("text", x = 4.75, y = .8, label = "Cases with negative residuals", color = "blue3") +
  labs(x = expression(paste("NEGEMOT: Negative emotions about climate change (", italic("X"), ")")),
       y = expression(paste("GOVACT: Support for governmentaction (", italic("Y"), ")"))) +
  coord_cartesian(xlim = range(glbwarm$negemot)) +
  theme_bw()
```

### Interpretation of the constant and regression coefficient.

Nothing to recode, here.

### The standardized regression model.

Same here.

### Simple linear regression with a dichotomous antecedent variable.

Here we add `sex` to the model.

```{r model3, cache = T, message = F, warning = F}
model3 <- 
  brm(data = glbwarm, 
      family = gaussian,
      govact ~ 1 + sex,
      cores = 4)
```

```{r}
print(model3)
```

Our model output is very close to that in the text. If you just wanted the coefficients, you might use the `fixef()` function.

```{r}
fixef(model3) %>% round(digits = 3)
```

Though not necessary, we used the `round()` function to reduce the number of significant digits in the output. 

You can get a little more information with the `posterior_summary()` function.

But since Bayesian estimation yields an entire posterior distribution, you can visualize that distribution in any number of ways. Because we'll be using ggplot2, we'll need to put the posterior draws into a data frame before plotting.

```{r}
post <- posterior_samples(model3)
```

We could summarize the posterior with boxplots

```{r, fig.width = 3, fig.height = 3}
post %>% 
  rename(female = b_Intercept) %>% 
  mutate(male = female + b_sex) %>% 
  select(male, female) %>% 
  gather() %>% 
  
  ggplot(aes(x = key, y = value)) +
  geom_boxplot(aes(fill = key)) +
  theme_bw() +
  theme(legend.position = "none")
```

or with overlapping density plots

```{r, fig.width = 4, fig.height = 2}
post %>% 
  rename(female = b_Intercept) %>% 
  mutate(male = female + b_sex) %>% 
  select(male, female) %>% 
  gather() %>% 
  
  ggplot(aes(x = value, group = key, fill = key)) +
  geom_density(color = "transparent", alpha = 3/4) +
  theme_bw()
```

or even with violin plots with superimposed posterior medians and 95% intervals.

```{r, fig.width = 3, fig.height = 2.5}
post %>% 
  rename(female = b_Intercept) %>% 
  mutate(male = female + b_sex) %>% 
  select(male, female) %>% 
  gather() %>% 
  
  ggplot(aes(x = key, y = value)) +
  geom_violin(aes(fill = key), color = "transparent", alpha = 3/4) +
  stat_summary(fun.y = median,
               fun.ymin = function(i){quantile(i, probs = .025)},
               fun.ymax = function(i){quantile(i, probs = .975)}) +
  theme_bw() +
  theme(legend.position = "none")
```

For even more ideas, see [Matthew Kay](https://twitter.com/mjskay)'s [tidybayes package](https://mjskay.github.io/tidybayes/articles/tidybayes.html). You can also get a sense of the model estimates for women and men with a little addition. Here we continue to use the `round()` function to simplify the output.

```{r}
# for women
round(fixef(model3)[1, ], digits = 2)

# for men
round(fixef(model3)[1, ] + fixef(model3)[2, ], digits = 2)
```

Here's the partially-standardized model, first with `lm()`.

```{r, message = F, warning = F}
glbwarm <-
  glbwarm %>% 
  mutate(govact_z = (govact - mean(govact))/sd(govact))

lm(data = glbwarm, govact_z ~ 1 + sex)
```

Now we'll use Bayes.

```{r model3_p_z, cache = T, message = F, warning = F}
model3_p_z <- 
  brm(data = glbwarm, 
      family = gaussian,
      govact_z ~ 1 + sex,
      cores = 4)
```

```{r}
fixef(model3_p_z)
```

## Alternative explanations for association

On page 46, Hayes produced a couple correlations. Here's how to get them from base R.

```{r}
cor(glbwarm$sex, glbwarm$negemot)
cor(glbwarm$sex, glbwarm$govact)
```

Again, if we wanted to get full Bayesian estimates, we'd fit an intercept-only multivariate model.

```{r, message = F, warning = F}
brm(data = glbwarm, 
    family = gaussian,
    mvbind(negemot, govact, sex) ~ 1,
    cores = 4) %>% 
  
  print(digits = 3)
```

For our purposes, the action is in the 'rescor($i$, $j$)' portions of the 'Family Specific Parameters' section.

Anyway, if you wanted to get all the Pearson's correlations among the `glbwarm` variables, rather than piecewise `cor()` approach, you could use the `lowerCor()` function from the [psych package](https://cran.r-project.org/web/packages/psych/index.html).

```{r}
psych::lowerCor(glbwarm[, 1:7], digits = 3)
```

## Multiple linear regression

here’s nothing particularly special about jumping from univariable to multivariable models in brms. You just keep tacking on predictors with the `+` operator.

```{r model4, cache = T, message = F, warning = F}
model4 <- 
  brm(data = glbwarm, 
      family = gaussian,
      govact ~ 1 + negemot + posemot + ideology + sex + age,
      cores = 4)
```

```{r}
print(model4)
```

Here is the posterior mean, what you might call the Bayesian point estimate, for someone with:

* negative emotions = 3,
* positive emotions = 4,
* `ideology` = 2,
* is male (i.e., `sex` = 1), and
* is 30 years of `age`.

```{r}
fixef(model4)[1] + 
  fixef(model4)[2] * 3 + 
  fixef(model4)[3] * 4 + 
  fixef(model4)[4] * 2 + 
  fixef(model4)[5] * 1 + 
  fixef(model4)[6] * 30
```

Here's the same deal for a man of the same profile, but with one point higher on `negemot`.

```{r}
fixef(model4)[1] + 
  fixef(model4)[2] * 4 + 
  fixef(model4)[3] * 4 + 
  fixef(model4)[4] * 2 + 
  fixef(model4)[5] * 1 + 
  fixef(model4)[6] * 30
```

If you want a full expression of the model uncertaintly in terms of the shape of the posterior distribution and the 95% intervals, you'll probably just want to use `posterior_samples()` and do a little data processing.

```{r, fig.width = 4, fig.height = 3}
post <- posterior_samples(model4)

post <-
  post %>%
  mutate(our_posterior = b_Intercept + b_negemot*4 + b_posemot*4 + b_ideology*2 + b_sex*1 + b_age*30)

# this intermediary step will make it easier to specify the break points and their labels for the x-axis  
post_summary <-
  quantile(post$our_posterior, probs = c(.025, .5, .975)) %>% 
  as_tibble() %>% 
  mutate(labels = value %>% 
           round(digits = 3) %>% 
           as.character())

# plot!
ggplot(data = post,
       aes(x = our_posterior)) +
  geom_density(fill = "black") +
  geom_vline(xintercept = post_summary$value,
             size = c(.5, .75, .5), linetype = c(2, 1, 2), color = "white") +
  scale_x_continuous(NULL,
                     breaks = post_summary$value,
                     labels = post_summary$labels) +
  scale_y_continuous(NULL, breaks = NULL) +
  labs(subtitle = "The expected govact score for a 30-year-old man for\nwhom negemot and posemot both equal 4 and ideology\nequals 2. The solid and dashed white vertical lines are the\nposterior median and 95% intervals, respectively.") +
  theme_bw()
```

In the text, Hayes showed that individuals based on these two profiles would be expected to differ by 0.441 (i.e., 5.244 - 4.803 = 0.441). That's fine if you're only working with OLS point estimates. But a proper Bayesian approach would express the difference in terms of an entire poster distribution, or at least a point estimate accompanied by some sort of intervals. Here we'll just work with the posterior to create a difference distribution. You could do that with a little deft `posterior_samples()` wrangling. Here we'll employ `fitted()`.

```{r, fig.width = 4, fig.height = 3}
nd <-
  tibble(negemot  = c(3, 4),
         posemot  = 4,
         ideology = 2,
         sex      = 1,
         age      = 30)

fitted(model4, 
         newdata = nd,
         summary = F) %>% 
  data.frame() %>% 
  set_names(str_c("condition_", letters[1:2])) %>% 
  mutate(difference = condition_b - condition_a) %>% 
  
  ggplot(aes(x = difference)) +
  geom_density(fill = "black", color = "transparent") +
  scale_y_continuous(NULL, breaks = NULL) +
  ggtitle("The posterior density for the difference between\nthe two conditions.") +
  theme_bw()
```

### The standardized regression model.

brms doesn't automatically give us the standardized coefficients the way OLS output often does, we'll have to be proactive. One solution is just to standardized the data themselves and then re-fit the model with those standardized variables. That leads us to the issue of how one standardized variables to begin with. Recall that standardizing entails subtracting the mean of a variable from that variable and then dividing that value by the standard deviation. We don't want to do that by hand. So one handy way is to make a custom function to do that work for us.

```{r}
sandardize <- function(x){
  (x - mean(x)) / sd(x)
}
```

Go [here](http://r4ds.had.co.nz/functions.html) to learn more about making custom functions in R.

Here we'll employ our custom `standardize()` function to make standardized versions of our variables.

```{r}
glbwarm <-
  glbwarm %>% 
  mutate(posemot_z  = sandardize(posemot), 
         negemot_z  = sandardize(negemot), 
         ideology_z = sandardize(ideology),
         sex_z      = sandardize(sex),
         age_z      = sandardize(age))
```

Now we've got us our standardized variables, let's fit a standardized model.

```{r model4_z, cache = T, message = F, warning = F}
model4_z <- 
  brm(data = glbwarm, 
      family = gaussian,
      govact_z ~ 1 + negemot_z + posemot_z + ideology_z + sex_z + age_z,
      cores = 4)
```

Here are the newly standardized coefficient summaries, minus the `Intercept`.

```{r}
fixef(model4_z)[-1, ] %>% round(3)
```

Our coefficients match up nicely with those in the text. Just as with Hayes's OLS estimates, we should not attempt to interpret the standardized `sex_z` coefficient from our Bayesian model.

Here's how we'd fit a *partially*-standardized model--a model in which all variables except for `sex` are standardized.

```{r model4_z_p, cache = T, message = F, warning = F}
model4_z_p <- 
  update(model4_z,
         newdata = glbwarm,
         formula = govact_z ~ 1 + negemot_z + posemot_z + ideology_z + sex + age_z,
         cores = 4)
```


Notice our use of the `update()` function. If you want to hastily fit a **brms** model of the same basic form of a prior model, you can just update some of the parameters of that original fit. In this case, all we did was swap out one of the predictors. To accommodate that change, we used the `newdata` argument so the model had access to the new variable and then we fed in the new formula. 

Anyway, here are the coefficient summaries, including the `Intercept`, for the partially-standardized model.

```{r}
fixef(model4_z_p) %>% round(digits = 3)
```

As Hayes wrote, now `sex` = `r round(brms::fixef(model4_z_p)[5], digits = 3)` has a sensible interpretation. "We can say that men and women differ by [`r round(brms::fixef(model4_z_p)[5], digits = 3)`] standard deviations in their support for government action when all other variables in the model are held constant (p. 53)."

On page 54, Hayes gave us the equation to transform unstandardized coefficients to standardized ones:

$$
\tilde{b}_{i} = b_{i}\left(\frac{SD_{X_{i}}}{SD_{Y}}\right)
$$

Let's give it a whirl with `negemot`.

```{r}
# here's the coefficient for `negemot` from the standardized model, `model4_z`
fixef(model4_z)["negemot_z", "Estimate"]

# here's the coefficient for `negemot` from the unstandardized model, `model4`
fixef(model4)["negemot", "Estimate"]

# and here we use Hayes' formula to standardize the unstandardized coefficient
fixef(model4)["negemot", "Estimate"] * (sd(glbwarm$negemot) / sd(glbwarm$govact))
```

Looks like we got it within rounding error--pretty good! However, that was just the posterior mean, the Bayesian point estimate. If we want to more fully express the uncertainty around the mean--and we do--, we'll need to work with the posterior draws.

```{r}
# the posterior draws from the unstandardized model
posterior_samples(model4) %>% 
  # using Hayes' formula to standardize `b_negemot`
  mutate(`hand-made b_negemot_z` = b_negemot * (sd(glbwarm$negemot) / sd(glbwarm$govact))) %>% 
  # taking on the `b_negemot_z` column from the standardized `model4_z` models posterior draws
  bind_cols(posterior_samples(model4_z) %>%
              select(b_negemot_z)) %>% 
  # isolating those two columns
  select(`hand-made b_negemot_z`, b_negemot_z) %>% 
  # converting the data to the long format and grouping by `key`
  gather() %>% 
  group_by(key) %>% 
  # here we summarize the results
  summarise(mean = mean(value),
            sd = sd(value),
            ll = quantile(value, probs = .025),
            ul = quantile(value, probs = .975)) %>% 
  mutate_if(is.double, round, digits = 3)
```

Our summary confirms that we can apply Hayes' formula to a `posterior_samples()` column in order to get fuller summary statistics for a hand-converted standardized coefficient. This would be in full compliance with, say, APA recommendations to include 95% intervals with all effect sizes--the standardized regression coefficient being the effect size, here.

## Measures of model fit

In the Bayesian world, we don't tend to appeal to the $SS_{residual}$, the $MS_{residual}$, or the standard error of estimate. We do sometimes, however, appeal to the $R^2$. I'm not going to go into the technical details here, but you should be aware that the Bayesian $R^2$ is not calculated the same as the OLS $R^2$. If you want to dive in, check out the paper by [Gelman, Goodrich, Gabry, and Ali](https://github.com/jgabry/bayes_R2/blob/master/bayes_R2.pdf). Here's how to get it in brms.

```{r}
bayes_R2(model4, summary = T) %>% round(digits = 3)
```

Happily, it comes with 95% intervals, which will make the editors at APA journals happy. If you want to go beyond summary statistics and take a look at the full posterior, just set `summary = F` and data wrangle and plot as usual.

```{r, fig.width = 6, fig.height = 1.75}
bayes_R2(model4, summary = F) %>% 
  data.frame() %>% 
  
  ggplot(aes(x = R2)) +
  geom_density(fill = "black", color = "transparent") +
  scale_y_continuous(NULL, breaks = NULL) +
  labs(title = expression(paste("Behold: The Bayesian ", italic("R")^{2}, " distribution for model 4")),
       x = NULL) +
  coord_cartesian(xlim = 0:1) +
  theme_bw()
```

Another way we examine model fit is with graphical posterior predictive checks. Posterior predictive checking is a very general approach, which you might learn more about [here](http://www.stat.columbia.edu/~gelman/research/published/bayes-vis.pdf) or with a few keyword searches in on [Gelman's blog](http://andrewgelman.com). One basic way is to use the model in order to simulate data and then compare those data with the original data--the basic idea being that good fitting models should produce data similar to the original data.

Recall how we've used `fitted()` to make regression lines and expected values? We'll, now we'll use `predict()` to simulate data based on our models.

```{r, fig.width = 10, fig.height = 3}
predict(model4, 
        summary = F,
        nsamples = 3) %>%
  t() %>% 
  as_tibble() %>% 
  gather() %>% 
  mutate(key = str_replace(key, "V", "simulation ")) %>% 
  rename(govact = value) %>% 
  bind_cols(
    bind_rows(
      glbwarm %>% select(-govact),
      glbwarm %>% select(-govact),
      glbwarm %>% select(-govact))
    ) %>% 
  
  ggplot(aes(x = negemot, y = govact)) +
  geom_jitter(height = .05, width = .05, 
              alpha = 1/2, size = 1/3) +
  coord_cartesian(ylim = 0:9) +
  theme_bw() +
  facet_wrap(~key, ncol = 3)
```

The question is, do these simulated data sets look like the original data? Let's see.

```{r, fig.width = 3.35, fig.height = 2.7}
glbwarm %>% 
  ggplot(aes(x = negemot, y = govact)) +
  geom_jitter(height = .05, width = .05, 
              alpha = 1/2, size = 1/3) +
  coord_cartesian(ylim = 0:9) +
  theme_bw()
```

Overall, the simulations aren't bad. But in all three `govact` tends to veer above 7.5, which is where the original data appear to be bounded. But otherwise the overall shape is pretty close, at least with respect to `negemot`. 

There's nothing special about three simulations. Three is just more than one and gives you a sense of the variance across simulations. Also, we only examined the model fit with respect to `negemot`. Since there are other variables in the model, we might also assess the model based on them.

Another method is with the `brms::pp_check()` function, which allows users to access a variety of convenience functions from the [bayesplot package](https://github.com/stan-dev/bayesplot). Here we'll use the default settings and just tack on `theme_bw()` for aesthetics.

```{r, fig.width = 6, fig.height = 3}
pp_check(model4) +
  theme_bw()
```

What we did was simulate 10 data sets worth of `govact` values, plot their densities (i.e., the thin blue lines) and compare them with the density of the original `govact` values. What we want is for the thin blue lines to largely align with the thick blue line. Though not perfect, the simulations from our `model4` did a pretty okay job of reproducing the original `govact` distribution. For more ideas on this method, see the brms [reference manual](https://cran.r-project.org/web/packages/brms/brms.pdf) and [this vignette](https://cran.r-project.org/web/packages/bayesplot/vignettes/graphical-ppcs.html) for the bayesplot package.

## Statistical inference

Here's a tidyverse way to do Hayes' simulation. We're just using OLS regression with the `lm()` function. You could do this with Bayesian HMC estimation, but man would it take a while. For our first step, we’ll define two custom functions.

```{r}
# this first one will use the `sample_n()` function to randomly sample from `glbwarm`
make_sample <- function(i) {
  set.seed(i)
  sample_n(glbwarm, 50, replace = F)
}

# this second function will fit our model, the same one from `model4`, to each of our subsamples
glbwarm_model <- function(df) {
  lm(govact ~ 1 + negemot + posemot + ideology + sex + age, data = df)
}
```

Now we'll run the simulation, wrangle the output, and plot in one fell swoop.

```{r sim, cache = T, fig.width = 6, fig.height = 3.5}
# we need an iteration index, which will double as the values we set our seed with in our `make_sample()` function
tibble(iter = 1:1e4) %>% 
  group_by(iter) %>% 
  # inserting our subsamples
  mutate(sample = map(iter, make_sample)) %>% 
  # fitting our models
  mutate(model = map(sample, glbwarm_model)) %>% 
  # taking those model results and tidying them with the broom package
  mutate(broom = map(model, broom::tidy)) %>% 
  # unnesting allows us to access our model results
  unnest(broom) %>% 
  # we're only going to focus on the estimates for `negemot`
  filter(term == "negemot") %>% 
  
  # here it is, Figure 2.7
  ggplot(aes(x = estimate)) +
  geom_histogram(binwidth = .025, boundary = 0) +
  labs(x = "Unstandardized regression coefficient for negemot",
       y = "Frequency in 1e4 samples of size 50") +
  theme_bw()
```

To learn more about this approach to simulations, see [this section of *R4DS*](http://r4ds.had.co.nz/many-models.html#nested-data).

### Testing a null hypothesis.

As Bayesians, we don't need to wed ourselves to the null hypothesis. We're not interested in the probability of the data given the null hypothesis. Bayes' rule,

$$p(\theta | D) = \frac{p(D | \theta) p(\theta)}{p(D)}$$

gives us the probability of the model parameters, given the data. Though I acknowledge [different researchers might set out to ask different things of their data](http://daniellakens.blogspot.com/2017/11/the-statisticians-fallacy.html), I propose we're generally more interested in determining the most probable parameter values than we are the probabilities tested within the NHST paradigm.

```{r, fig.width = 6, fig.height = 3.5}
posterior_samples(model4) %>% 
  
  ggplot(aes(x = b_negemot)) +
  geom_density(fill = "black", color = "transparent") +
  geom_vline(xintercept = posterior_interval(model4)["b_negemot", ],
             color = "white", linetype = 2) +
  scale_x_continuous(breaks = posterior_interval(model4)["b_negemot", ] %>% as.double(),
                     labels = posterior_interval(model4)["b_negemot", ] %>% as.double() %>% round(digits = 2) %>% as.character()) +
  scale_y_continuous(NULL, breaks = NULL) +
  labs(subtitle = "The most probable values for our b_negemot parameter are the ones around the peak\nof the density. For convenience, the dashed lines denote the 95% credible intervals.\nSure, you could ask yourself, 'Is zero within those intervals?' But with such rich output,\nthat seems like an impoverished question to ask.") +
  theme_bw()
```

### Interval estimation.

Within the Bayesian paradigm, we don't use 95% intervals based on the typical frequentist formula. With the brms package, we typically use percentile-based intervals. Take the 95% credible intervals for the `negemot` coefficient from model `model4`:

```{r}
posterior_interval(model4)["b_negemot", ]
```

We can actually get those intervals with the simple use of the base R `quantile()` function.

```{r}
posterior_samples(model4) %>% 
  summarize(the_2.5_percentile  = quantile(b_negemot, probs = .025),
            the_97.5_percentile = quantile(b_negemot, probs = .975))
```

The consequence of this is that our Bayesian credible intervals aren't necessarily symmetric. Which is fine because the posterior distribution for a given parameter isn't always symmetric.

But not all Bayesian intervals are percentile based. John Kruschke, for example, often recommends highest posterior density intervals in [his work](http://www.indiana.edu/~kruschke/DoingBayesianDataAnalysis/). The brms package doesn't have a convenience function for these, but you can compute them with help from the [HDInterval package](https://cran.r-project.org/web/packages/HDInterval/index.html). 

```{r}
library(HDInterval)

hdi(posterior_samples(model4)[ , "b_negemot"], credMass = .95)
```

Finally, because Bayesians aren't bound to the NHST paradigm, we aren’t bound to 95% intervals, either. For example, in both his [excellent text](http://xcelab.net/rm/statistical-rethinking/) and as a default in its accompanying [rethinking package](https://github.com/rmcelreath/rethinking), Richard McElreath often uses 89% intervals. Alternatively, Andrew Gelman has publically advocated for [50% intervals](http://andrewgelman.com/2016/11/05/why-i-prefer-50-to-95-intervals/). The most important thing is to express the uncertainty in the posterior in a clearly-specified way. If you'd like, say, 80% intervals in your model summary, you can insert a `prob` argument into either `print()` or `summary()`.

```{r}
print(model4, prob = .8)
```

Note how two of our columns changed to 'l-80% CI' and 'u-80% CI'.

You can specify custom percentile levels with `posterior_summary()`:

```{r}
posterior_summary(model4, probs = c(.9, .8, .7, .6, .4, .3, .2, .1)) %>% 
  round(digits = 2)
```

And of course, you can use multiple interval summaries when you `summarize()` the output from `posterior_samples()`. E.g.,

```{r}
posterior_samples(model4) %>% 
  select(b_Intercept:b_age) %>% 
  gather() %>% 
  group_by(key) %>% 
  summarize(`l-95% CI` = quantile(value, probs = .025),
            `u-95% CI` = quantile(value, probs = .975),
            `l-50% CI` = quantile(value, probs = .25),
            `u-50% CI` = quantile(value, probs = .75)) %>% 
  mutate_if(is.double, round, digits = 2)
```

Throughout this project, I'm going to be lazy and default to conventional 95% intervals, with occasional appearances of 50% intervals.

### Testing a hypothesis about a set of antecedent variables.

Here we'll make use of the `update()` function to hastily fit our reduced model, `model5`.

```{r model5, cache = T, message = F, warning = F}
model5 <- 
  update(model4,
         govact ~ 1 + ideology + sex + age,
         cores = 4)
```

We can get a look at the $R^2$ summaries for our competing models like this.

```{r}
bayes_R2(model4) %>% round(digits = 2)
bayes_R2(model5) %>% round(digits = 2)
```

So far, it looks like our fuller model, `model4`, explains more variation in the data. If we wanted to look at their distributions, we'd set `summary = F` in the `bayes_R2()` function and convert the results to a data frame or tibble. Here we use `bind_cols()` to put the $R^2$ results for both in the same tibble.

```{r}
r2s <-
  bayes_R2(model4, summary = F) %>%
  data.frame() %>%
  rename(r2_model4 = R2) %>% 
  bind_cols(
    bayes_R2(model5, summary = F) %>%
      data.frame() %>%
      rename(r2_model5 = R2)
  )

head(r2s)
```

With our `r2s` tibble in hand, we're ready to plot.

```{r, fig.width = 6, fig.height = 1.5}  
r2s %>% 
  gather() %>% 
  
  ggplot(aes(x = value, fill = key)) +
  geom_density(color = "transparent") +
  scale_y_continuous(NULL, breaks = NULL) +
  coord_cartesian(xlim = 0:1) +
  theme_bw()
```

Yep, the $R^2$ distribution for `model4`, the one including the emotion variables, is clearly larger than that for the more parsimonious `model5`. And it'd just take a little more data wrangling to get a formal $R^2$ difference score.

```{r, fig.width = 6, fig.height = 2}  
r2s %>% 
  mutate(difference = r2_model5 - r2_model4) %>% 
  
  ggplot(aes(x = difference)) +
  geom_density(fill = "black", color = "transparent") +
  scale_y_continuous(NULL, breaks = NULL) +
  coord_cartesian(xlim = -1:0) +
  labs(title = expression(paste(Delta, italic("R")^{2})),
       subtitle = expression(paste("This is the amount the ", italic("R")^{2}, " dropped after pruning the emotion variables from the model.")),
       x = NULL) +
  theme_bw()
```

The $R^2$ approach is popular within the social sciences. But it has its limitations, the first of which is that it doesn't correct for model complexity. The second is it’s not applicable to a range of models, such as those that do not use the Gaussian likelihood (e.g., logistic regression) or to multilevel models.

Happily, information criteria offer a more general framework. The AIC is the most popular information criteria among frequentists. Within the Bayesian world, we have the DIC, the WAIC, and the LOO. The DIC is quickly falling out of favor and is not immediately available with the brms package. However, we can use the WAIC and the LOO, both of which are computed in brms via the [loo package](https://github.com/stan-dev/loo).

In brms, you can get the WAIC or LOO values with `waic()` or `loo()`, respectively. Here we use `loo()` and save the output as objects.

```{r models_4_and_5_loos, cache = T, message = F, warning = F}
l_model4 <- loo(model4)
l_model5 <- loo(model5)
```

Here's the main loo-summary output for `model4`.

```{r}
l_model4
```

You get a wealth of output, more of which can be seen with `str(l_model4)`. For now, notice the "All pareto k estimates are good (k < 0.5)." Pareto $k$ values [can be used for diagnostics](https://cran.r-project.org/web/packages/loo/vignettes/loo2-example.html#plotting-pareto-k-diagnostics). Each case in the data gets its own $k$ value and we like it when those $k$s are low. The makers of the loo package get worried when those $k$s exceed 0.7 and as a result, `loo()` spits out a warning message when they do. Happily, we have no such warning messages in this example.

If you want to work with the $k$ values directly, you can extract them and place them into a tibble like so:

```{r}
l_model4$diagnostics %>% 
  as_tibble()
```

The `pareto_k` values can be used to examine cases that are overly-influential on the model parameters, something akin to a Cook's $D_{i}$. See, for example [this discussion on stackoverflow.com](https://stackoverflow.com/questions/39578834/linear-model-diagnostics-for-bayesian-models-using-rstan/39595436) in which several members of the [Stan team](http://mc-stan.org) weighed in. The issue is also discussed in [this paper](https://arxiv.org/abs/1507.04544) and in [this lecture by Aki Vehtari](https://www.youtube.com/watch?v=FUROJM3u5HQ&feature=youtu.be&a=).

But anyway, we're getting ahead of ourselves. Back to the LOO.

Like other information criteria, the LOO values aren't of interest in and of themselves. However, the values of one model's LOO relative to that of another is of great interest. We generally prefer models with lower information criteria. With `compare_ic()`, we can compute a formal difference score between multiple loo objects.

```{r}
compare_ic(l_model4, l_model5)
```

We also get a standard error. Here it looks like `model4` was substantially better, in terms of LOO-values, than `model5`.

For more on the LOO, see the [loo reference manual](https://cran.r-project.org/web/packages/loo/loo.pdf), this [handy vignette](https://cran.r-project.org/web/packages/loo/vignettes/loo2-example.html), or the scholarly papers referenced therein. Also, although McElreath doesn't discuss the LOO, he does cover the topic in general in his [text](http://xcelab.net/rm/statistical-rethinking/) and in his [online lectures on the topic](https://www.youtube.com/watch?v=t0pRuy1_190&t=1307s&list=PLDcUM9US4XdM9_N6XUUFrhghGJ4K25bFc&index=9).

## Multicategorical antecedent variables

`partyid` is coded 1 = Democrat 2 = Independent 3 = Repuclican. You can get a count of the cases within a give `partyid` like this:

```{r}
glbwarm %>% 
  group_by(partyid) %>% 
  count()
```

We can get grouped means for `govact` like this:

```{r}
glbwarm %>% 
  group_by(partyid) %>% 
  summarize(mean_support_for_governmental_action = mean(govact))
```

We can make dummies with the `ifelse()` function. We'll just go ahead and do that right within the `brm()` function.

```{r model6, cache = T, message = F, warning = F}
model6 <- 
  brm(data = glbwarm %>% 
        mutate(Democrat = ifelse(partyid == 1, 1, 0),
               Republican = ifelse(partyid == 3, 1, 0)), 
      family = gaussian,
      govact ~ 1 + Democrat + Republican,
      chains = 4, cores = 4)
```

```{r}
fixef(model6)
```

The intercept is the stand-in for Independents and the other two coefficients are difference scores.

The $R^2$ is okay:

```{r}
bayes_R2(model6) %>% round(digits = 3)
```

There's no need to compute an $F$-test on our $R^2$. The posterior mean and it's 95% intervals are well away from zero. But you could use your `bayes_R2(model6, summary = F)` plotting skills from above to more fully inspect the posterior if you'd like.

We could also use information criteria. One method would be to compare the WAIC or LOO value of `model6` with an intercept-only model. First, we'll need to fit that model.

```{r model7, cache = T, message = F, warning = F}
model7 <- 
  update(model6,
         govact ~ 1,
         chains = 4, cores = 4)
```

If we just put both models into `waic()`, we can side step the need to save their outputs as objects which we then put into `compare_ic()`.

```{r models_6_and_7_waics, cache = T, message = F, warning = F}
waic(model6, model7)
```

The WAIC comparison suggests `model6`, the one with the `partyid` dummies, is an improvement over the simple intercept-only model. Another way to compare the information criteria is with AIC-type weighting. The brms package offers a variety of weighting methods via the `model_weights()` function.

```{r model_weights, cache = T, message = F, warning = F}
MWs <- model_weights(model6, model7, weights = "waic")

MWs
```

If you're not a fan of scientific notation, you can put the results in a tibble and look at them on a plot.

```{r, fig.width = 5, fig.height = 1.5}
MWs %>% 
  as.data.frame() %>% 
  rownames_to_column() %>% 

  ggplot(aes(x = ., y = rowname)) +
  geom_point() +
  labs(subtitle = "The weights should sum to 1. In this case virtually all the weight is placed\nin model6. Recall, that these are RELATIVE weights. Add another\nmodel fit into the mix and the weights might well change.", 
       x = "WAIC weight", y = NULL) +
  theme_bw() +
  theme(axis.ticks.y = element_blank())
```

You could, of course, do all this with the LOO.

## Assumptions for interpretation and statistical inference

### Linearity. 

I have nothing to add, here.

### Normality.

The brms package is quite general and allows users to fit models from a variety of likelihoods other than the Gaussian. For example, users can accommodate outliers/extreme values with [Student's t regression](https://github.com/ASKurz/Student-s-t_regression). You can do count regression with the Poisson or the negative binomial... For more, see McElreath's lecture [introducing the generalized linear model](https://www.youtube.com/watch?v=DyrUkqK9Tj4&t=254s&frags=pl%2Cwn) or Bürkner's vignette, [*Parameterization of Response Distributions in brms*](https://cran.r-project.org/web/packages/brms/vignettes/brms_families.html)

### Homoscedasticity.

The brms package can also accommodate homoscedasticity with distributional modeling. In short, one simply models $\sigma$ in addition to the mean, $\mu$. See Bürkner's [handy vignette](https://cran.r-project.org/web/packages/brms/vignettes/brms_distreg.html) on the topic.

### Independence.

And the issue of independence is where the multilevel model comes on. See any relevant text, such as [*Statistical Rethinking*](http://xcelab.net/rm/statistical-rethinking/) or [*Data Analysis Using Regression and Multilevel/Hierarchical Models*](http://www.stat.columbia.edu/~gelman/arm/).

## Reference {-}

[Hayes, A. F. (2018). *Introduction to mediation, moderation, and conditional process analysis: A regression-based approach.* (2nd ed.). New York, NY, US: The Guilford Press.](http://afhayes.com/introduction-to-mediation-moderation-and-conditional-process-analysis.html)

## Session info {-}

```{r}
sessionInfo()
```
