# More Than One Mediator

## The parallel multiple mediator model

```{r set-options, echo = FALSE, cache = FALSE}
options(width = 110)
```

## Example using the presumed media influence study

Here we load a couple necessary packages, load the data, and take a `glimpse()`.

```{r, warning = F, message = F}
library(tidyverse)

pmi <- read_csv("data/pmi/pmi.csv")

glimpse(pmi)
```

Let's load brms.

```{r, message = F, warning = F}
library(brms)
```

Bayesian correlations, recall, just take an intercepts-only multivariate model.

```{r model1, cache = T, message = F, warning = F}
model1 <- 
  brm(data = pmi, family = gaussian,
      cbind(pmi, import) ~ 1,
      chains = 4, cores = 4)
```

A little indexing with the `posterior_summary()` function will get us the Bayesian correlation with its posterior $SD$ and intervals.

```{r}
posterior_summary(model1)["rescor__pmi__import", ] %>% round(digits = 3)
```

As with single mediation models, the multiple mediation model requires we carefully construct the formula for each criterion. We'll continue to use the multiple `bf()` approach.

```{r}
m1_model <- bf(import   ~ 1 + cond)
m2_model <- bf(pmi      ~ 1 + cond)
y_model  <- bf(reaction ~ 1 + import + pmi + cond)
```

And now we fit the model.

```{r model2, cache = T, message = F, warning = F}
model2 <-
  brm(data = pmi, family = gaussian,
      y_model + m1_model + m2_model + set_rescor(FALSE),
      chains = 4, cores = 4)
```

```{r}
print(model2, digits = 3)
```

Because we have three criteria, we'll have three Bayesian $R^2$ posteriors.

```{r, fig.width = 8, fig.height = 2}
library(ggthemes)

bayes_R2(model2, summary = F) %>% 
  as_tibble() %>% 
  gather() %>% 
  mutate(key = str_remove(key, "R2_")) %>% 
  
  ggplot(aes(x = value, color = key, fill = key)) +
  geom_density(alpha = .5) +
  scale_color_ptol() +
  scale_fill_ptol() +
  scale_y_continuous(NULL, breaks = NULL) +
  coord_cartesian(xlim = 0:1) +
  labs(title = expression(paste("Our ", italic("R")^{2}, " distributions")),
       subtitle = "The densities for import and pmi are asymmetric, small, and largely overlapping. The density for reaction is Gaussian and\nmore impressive in magnitude.",
       x = NULL) +
  theme_minimal() +
  theme(legend.title = element_blank())
```

It'll take a bit of data wrangling to rename our model parameters to the $a$, $b$... configuration. We'll compute the indirect effects and $c$, too.

```{r}
post <- posterior_samples(model2)

post<-
  post %>% 
  mutate(a1 = b_import_cond,
         a2 = b_pmi_cond,
         b1 = b_reaction_import,
         b2 = b_reaction_pmi,
         c_prime = b_reaction_cond) %>% 
  mutate(a1b1 = a1*b1,
         a2b2 = a2*b2) %>% 
  mutate(c = c_prime + a1b1 + a2b2)
```

Here are their summaries. Since Bayesians use means, medians, and sometimes the mode to describe the central tendencies of a parameter, this time we'll mix it up and just use the median. 

```{r}
post %>% 
  select(a1:c) %>% 
  gather() %>% 
  group_by(key) %>% 
  summarize(median = median(value), 
            ll = quantile(value, probs = .025),
            ul = quantile(value, probs = .975)) %>% 
  mutate_if(is_double, round, digits = 3)
```

```{r}
post %>% 
  mutate(dif = a1b1*b1) %>% 
  summarize(median = median(dif), 
            ll = quantile(dif, probs = .025),
            ul = quantile(dif, probs = .975)) %>% 
  mutate_if(is_double, round, digits = 3)
```

In the middle paragraph of page 158, Hayes showd how the mean difference in `imprt` between the two `cond` groups multiplied by `b1`, the coefficient of `import` predicting `reaction`, is equal to the `a1b1` indirect effect. He does this with simple algebra using the group means and the point estimates. 

Let's follow along. First, we'll get those two group means and save them as numbers to arbitrary precision.

```{r}
(
  import_means <-
    pmi %>%
    group_by(cond) %>% 
    summarize(mean = mean(import))
 )

(cond_0_import_mean <- import_means[1, 2] %>% pull())
(cond_1_import_mean <- import_means[2, 2] %>% pull())
```

Here we follow the formula in the last sentence of the paragraph and then compare the results to the posterior for `a1b1`.

```{r}
post %>% 
  # Using his formula to make our new vector, `handmade a1b1` 
  mutate(`handmade a1b1` = (cond_1_import_mean - cond_0_import_mean)*b1) %>% 
  # Here's all the usual data wrangling
  select(a1b1, `handmade a1b1`) %>% 
  gather() %>% 
  group_by(key) %>% 
  summarize(mean = mean(value), 
            median = median(value), 
            ll = quantile(value, probs = .025),
            ul = quantile(value, probs = .975)) %>% 
  mutate_if(is_double, round, digits = 3)
```

Yep, at the mean, Hayes's formula is spot on. But the distributions are distinct. They differ slightly at the median and vastly in the widths of the posterior intervals. I’m no mathematician, so take this with a grain of salt, but I suspect this has to do with how we used fixed values (i.e., the difference of the subsample means) to compute `handmade a1b1`, but all the components in `a1b1` were estimated.

Here we'll follow the same protocol for `a2b2`.

```{r}
(
  pmi_means <-
    pmi %>%
    group_by(cond) %>% 
    summarize(mean = mean(pmi))
 )

cond_0_pmi_mean <- pmi_means[1, 2] %>% pull()
cond_1_pmi_mean <- pmi_means[2, 2] %>% pull()
```

```{r}
post %>% 
  mutate(`handmade a2b2` = (cond_1_pmi_mean - cond_0_pmi_mean)*b2) %>% 
  select(a2b2, `handmade a2b2`) %>% 
  gather() %>% 
  group_by(key) %>% 
  summarize(mean = mean(value), 
            median = median(value), 
            ll = quantile(value, probs = .025),
            ul = quantile(value, probs = .975)) %>% 
  mutate_if(is_double, round, digits = 3)
```

To get the total indirect effect as discussed on page 160, we simply add the `a1b1` and `a2b2` columns.

```{r}
post <-
  post %>% 
  mutate(total_indirect_effect = a1b1 + a2b2) 

post %>% 
  select(total_indirect_effect) %>% 
  summarize(mean = mean(total_indirect_effect), 
            median = median(total_indirect_effect), 
            ll = quantile(total_indirect_effect, probs = .025),
            ul = quantile(total_indirect_effect, probs = .975)) %>% 
  mutate_if(is_double, round, digits = 3)
```

To use the equations on the top of page 161, we'll just work directly with the original vectors in `post`.

```{r}
post %>% 
  mutate(Y_bar_given_X_1 = b_import_Intercept + b_reaction_cond*1 + b_reaction_import*b_import_Intercept + b_reaction_pmi*b_pmi_Intercept,
         Y_bar_given_X_0 = b_import_Intercept + b_reaction_cond*0 + b_reaction_import*b_import_Intercept + b_reaction_pmi*b_pmi_Intercept) %>% 
  mutate(`c_prime by hand` = Y_bar_given_X_1 - Y_bar_given_X_0) %>% 
  select(c_prime, `c_prime by hand`) %>% 
  gather() %>% 
  group_by(key) %>% 
  summarize(mean = mean(value), 
            median = median(value), 
            ll = quantile(value, probs = .025),
            ul = quantile(value, probs = .975))
```

We computed `c` a while ago.

```{r}
post %>% 
  summarize(mean = mean(c), 
            median = median(c), 
            ll = quantile(c, probs = .025),
            ul = quantile(c, probs = .975))
```

And `c` minus `c_prime` is straight subtraction.

```{r}
post %>% 
  mutate(`c minus c_prime` = c - c_prime) %>% 
  summarize(mean = mean(`c minus c_prime`), 
            median = median(`c minus c_prime`), 
            ll = quantile(`c minus c_prime`, probs = .025),
            ul = quantile(`c minus c_prime`, probs = .975))
```

## Statistical inference

### Inference about the direct and total effects.

We're not going to bother with $p$-values and we've already computed the 95% Bayesian credible intervals, above. But we can examine our parameters with a density plot.

```{r, fig.width = 8, fig.height = 2.5}
post %>% 
  select(c, c_prime) %>% 
  gather() %>%
  
  ggplot(aes(x = value, fill = key, color = key)) +
  geom_vline(xintercept = 0, color = "black") +
  geom_density(alpha = .5) +
  scale_color_ptol(NULL) +
  scale_fill_ptol(NULL) +
  scale_y_continuous(NULL, breaks = NULL) +
  labs(title = expression(paste("It appears zero is more credible for the direct effect, ", italic(c), "', than it is the total effect, ", italic(c), ".")),
       x = NULL) +
  coord_cartesian(xlim = -c(-1.5, 1.5)) +
  theme_minimal()
```

### Inference about specific indirect effects.

Again, no need to worry about bootstrapping within the Bayesian paradigm. We can compute high-quality percentile-based intervals with our HMC-based posterior samples.

```{r}
post %>%
  select(a1b1, a2b2) %>% 
  gather() %>% 
  group_by(key) %>% 
  summarize(ll = quantile(value, probs = .025),
            ul = quantile(value, probs = .975)) %>% 
  mutate_if(is.double, round, digits = 3)
```  

### Pairwise comparisons between specific indirect effects.

Within the Bayesian paradigm, it's straightforward to compare indirect effects. All one has to do is compute a difference score and summarize it somehow. Here it is, `a1b1` minus `a2b2`

```{r}
post <-
  post %>% 
  mutate(difference = a1b1 - a2b2) 

post %>%
  summarize(mean = mean(difference),
            ll = quantile(difference, probs = .025),
            ul = quantile(difference, probs = .975)) %>% 
  mutate_if(is.double, round, digits = 3)
```

Why not plot?

```{r, fig.width = 4.5, fig.height = 2.5}
post %>% 
  
  ggplot(aes(x = difference)) +
  geom_vline(xintercept = 0, color = "black", linetype = 2) +
  geom_density(color = "black", fill = "black", alpha = .5) +
  scale_y_continuous(NULL, breaks = NULL) +
  labs(title = "The difference score between the indirect effects",
       subtitle = expression(paste("No ", italic(p), "-value or 95% intervals needed for this one.")),
       x = NULL) +
  coord_cartesian(xlim = -1:1) +
  theme_minimal()
```

Although note well that this does not mean their difference is exactly zero. The shape of the posterior distribution testifies our uncertainty in their difference. Our best bet is that the difference is approximately zero, but it could easily be plus or minus a quarter of a point or more.

### Inference about the total indirect effect.

Here's the plot.

```{r, fig.width = 4.5, fig.height = 2.5}
post %>% 
  
  ggplot(aes(x = total_indirect_effect, fill = factor(0), color = factor(0))) +
  geom_density(alpha = .5) +
  scale_color_ptol() +
  scale_fill_ptol() +
  scale_y_continuous(NULL, breaks = NULL) +
  labs(title = "The total indirect effect of condition on reaction",
       subtitle = expression(paste("This is the sum of ", italic(a)[1], italic(b)[1], " and ", italic(a)[2], italic(b)[2], ". It's wide and uncertain.")),
       x = NULL) +
  theme_minimal() +
  theme(legend.position = "none")
```

## The serial multiple mediator model

### Direct and indirect effects in a serial multiple mediator model.

### Statistical inference.

### Example from the presumed media influence study.

The model syntax is similar to the earlier multiple mediator model. All we change is adding `import` to the list of predictors in the `m2_model`.

```{r}
m1_model <- bf(import   ~ 1 + cond)
m2_model <- bf(pmi      ~ 1 + import + cond)
y_model  <- bf(reaction ~ 1 + import + pmi + cond)
```

```{r model3, cache = T, message = F, warning = F}
model3 <-
  brm(data = pmi, family = gaussian,
      y_model + m1_model + m2_model + set_rescor(FALSE),
      chains = 4, cores = 4)
```

```{r}
print(model3)
```

The $R^2$ posterior densities:

```{r, fig.width = 8, fig.height = 2}
bayes_R2(model3, summary = F) %>% 
  as_tibble() %>% 
  gather() %>% 
  mutate(key = str_remove(key, "R2_")) %>% 
  
  ggplot(aes(x = value, color = key, fill = key)) +
  geom_density(alpha = .5) +
  scale_color_ptol() +
  scale_fill_ptol() +
  scale_y_continuous(NULL, breaks = NULL) +
  coord_cartesian(xlim = 0:1) +
  labs(title = expression(paste("The ", italic("R")^{2}, " distributions for model3, the serial multiple mediator model")),
       subtitle = "The density for reaction hasn't changed from model2. However, look how the pmi density separated from import.",
       x = NULL) +
  theme_minimal() +
  theme(legend.title = element_blank())
```

As before, here we'll save the posterior samples into a data frame and rename the parameters a bit to match Hayes's nomenclature.

```{r}
post <- posterior_samples(model3)

post <-
  post %>% 
  mutate(a1 = b_import_cond,
         a2 = b_pmi_cond,
         b1 = b_reaction_import,
         b2 = b_reaction_pmi,
         c_prime = b_reaction_cond,
         d21 = b_pmi_import)
```

Here are the parameter summaries for the pathways depicted in Figure 5.6.

```{r}
post %>% 
  select(a1:d21) %>% 
  gather() %>% 
  group_by(key) %>% 
  summarize(mean = mean(value), 
            median = median(value), 
            ll = quantile(value, probs = .025),
            ul = quantile(value, probs = .975)) %>% 
  mutate_if(is_double, round, digits = 3)
```

To get our version of the parameter summaries in Table 5.2, all you have to do is add the summaries for the intercepts to what we did above.

```{r}
post %>% 
  rename(im1 = b_import_Intercept,
         im2 = b_pmi_Intercept,
         iy = b_reaction_Intercept) %>% 
  select(a1:d21, starts_with("i")) %>% 
  gather() %>% 
  group_by(key) %>% 
  summarize(mean = mean(value), 
            median = median(value), 
            ll = quantile(value, probs = .025),
            ul = quantile(value, probs = .975)) %>% 
  mutate_if(is_double, round, digits = 3)
```

Here we compute the four indirect effects.

```{r}
post <-
  post %>%  
  mutate(a1b1 = a1*b1,
         a2b2 = a2*b2,
         a1d21b2 = a1*d21*b2) %>% 
  mutate(total_indirect_effect = a1b1 + a2b2 + a1d21b2)
```

The summaries for the indirect effects are as follows.

```{r}
post %>% 
  select(a1b1:total_indirect_effect) %>% 
  gather() %>% 
  group_by(key) %>% 
  summarize(mean = mean(value), 
            median = median(value), 
            ll = quantile(value, probs = .025),
            ul = quantile(value, probs = .975)) %>% 
  mutate_if(is_double, round, digits = 3)
```

To get the contrasts Hayes presented in page 179, we just do a little subtraction.

```{r}
post %>%  
  transmute(C1 = a1b1 - a2b2,
            C2 = a1b1 - a1d21b2,
            C3 = a2b2 - a1d21b2) %>% 
  gather() %>% 
  group_by(key) %>% 
  summarize(mean = mean(value), 
            median = median(value), 
            ll = quantile(value, probs = .025),
            ul = quantile(value, probs = .975)) %>% 
  mutate_if(is_double, round, digits = 3)
```

And just because it's fun, we may as well plot our indirect effects.

```{r, fig.width = 8, fig.height = 2.5}
# this will help us save a little space with the plot code
my_labels <- c(expression(paste(italic(a)[1], italic(b)[1])),
               expression(paste(italic(a)[1], italic(d)[21], italic(b)[1])),
               expression(paste(italic(a)[2], italic(b)[2])),
               "total indirect effect")
post %>% 
  select(a1b1:total_indirect_effect) %>% 
  gather() %>%
  
  ggplot(aes(x = value, fill = key, color = key)) +
  geom_density(alpha = .5) +
  scale_color_ptol(NULL, labels = my_labels,
                   guide = guide_legend(label.hjust = 0)) +
  scale_fill_ptol(NULL, labels = my_labels,
                   guide = guide_legend(label.hjust = 0)) +
  scale_y_continuous(NULL, breaks = NULL) +
  labs(title = "The four indirect effects of the serial multiple mediator model",
       x = NULL) +
  theme_minimal()
```

## References {-}

[Hayes, A. F. (2018). *Introduction to mediation, moderation, and conditional process analysis: A regression-based approach.* (2nd ed.). New York, NY, US: The Guilford Press.](http://afhayes.com/introduction-to-mediation-moderation-and-conditional-process-analysis.html)

## Session info {-}

```{r}
sessionInfo()
```
