---
title: "Chapter 03"
author: "A Solomon Kurz"
date: "`r format(Sys.Date())`"
output:
  github_document
---

```{r set-options, echo = FALSE, cache = FALSE}
options(width = 110)
```

## 3.2. Example with dichotomous $X$: The influence of presumed media influence

Here we load a couple necessary packages, load the data, and take a peek at them.

```{r, warning = F, message = F}
library(readr)
library(tidyverse)

p <- read_csv("data/pmi/pmi.csv")

glimpse(p)
```

You can get the male/female split like so:

```{r}
p %>% 
  group_by(gender) %>% 
  count()
```

Here is the split by `condition`:

```{r}
p %>% 
  group_by(cond) %>% 
  count()
```

Here is how to get the ungrouped Mean and $SD$ values for `reaction` and `pmi`, as presented in Table 3.1,

```{r}
p %>% 
  select(reaction, pmi) %>% 
  gather() %>% 
  group_by(key) %>% 
  summarise(mean = mean(value),
            sd = sd(value)) %>% 
  mutate_if(is.double, round, digits = 3)
```

You might get the Mean and $SD$ values for `reaction` and `pmi` grouped by `cond` like this:

```{r}
p %>% 
  select(reaction, pmi, cond) %>% 
  gather(key, value, -cond) %>% 
  group_by(cond, key) %>% 
  summarise(mean = mean(value),
            sd = sd(value)) %>% 
  mutate_if(is.double, round, digits = 3)
```

Let's load our primary statistical package.

```{r, message = F, warning = F}
library(brms)
```

Before we begin, I should acknowledge that I greatly benefited by [this great blog post](http://www.imachordata.com/bayesian-sem-with-brms/) by Jarrett Byrnes.

In brms, we handle mediation models using the [multivariate syntax](https://cran.r-project.org/web/packages/brms/vignettes/brms_multivariate.html). There are a few ways to do this. Let's start simple. 

If you look at the path model in Figure 3.3, you'll note that `reaction` is predicted by `pmi` and `cond`. `pmi`, in turn, is predicted solely by `cond`. So we have two regression models, which is just the kind of thing the brms multivariate syntax is for. So first, let's specify both models, which we'll nest in `bf()` functions and save as objects.

```{r}
y_model <- bf(reaction ~ 1 + pmi + cond)
m_model <- bf(pmi ~ 1 + cond)
```

Now we have our `bf()` objects in hand, we'll combine them with the `+` operator within the `brm()` function. We'll also specify `set_rescor(FALSE)`--we're not interested in adding a residual correlation between `reaction` and `pmi`.

```{r, message = F, warning = F}
fit0 <-
  brm(data = p, family = gaussian,
      y_model + m_model + set_rescor(FALSE),
      chains = 4, cores = 4)
```

Here are our results.

```{r}
print(fit0)
```

If you compare our model summary with the coefficients in the path model in Figure 3.3, you'll see our coefficients are the same. The brms summary also includes intercepts and residual variances, which are typically omitted in path diagrams, even though they're still part of the model.

If you're getting lost in all the model output, try this.

```{r}
fixef(fit0)[3:5, ] %>% round(digits = 3)
```

Also note that Hayes tends to refer to the intercepts as constants. 

In his Table 3.2, he also includes the $R^2$ values. Here are ours.

```{r}
bayes_R2(fit0) %>% round(digits = 3)
```

It's worth it to actually plot the $R^2$ distributions.

```{r, fig.width = 6, fig.height = 2}
library(ggthemes)

bayes_R2(fit0, summary = F) %>% 
  as_tibble() %>% 
  gather() %>% 
  
  ggplot(aes(x = value, fill = key)) +
  geom_density(color = "transparent", alpha = 2/3) +
  scale_fill_colorblind() +  # we got this color palette from the ggthemes package
  coord_cartesian(xlim = 0:1) +
  labs(title = expression(paste("The ", italic("R")^{2}, " distributions for fit0")),
       x = NULL) +
  theme_classic()
```

We went through the trouble of plotting the $R^2$ distributions because it’s useful to understand that they won’t often be symmetric when they’re near their logical boundaries (i.e., 0 and 1). This is where asymmetric Bayesian credible intervals can really shine.

Let's get down to business, the indirect effect, the $ab$ pathway. In our model

* $a$ = `pmi_cond`
* $b$ = `reaction_pmi`

You can isolate them with `fixef()[i]`.

```{r}
fixef(fit0)[5 , ]
fixef(fit0)[3 , ]
```

So the naive approach would be to just multiply them.

```{r}
(fixef(fit0)[5 , ] * fixef(fit0)[3 , ]) %>% round(digits = 3)
```

Now, this does get us the correct 'Estimate' (i.e., posterior mean). However, the posterior $SD$ and 95% intervals are off. If you want to do this properly, you need to work with the poster samples themselves. Here they are:

```{r}
post <- posterior_samples(fit0)

glimpse(post)
```

Here we compute the indirect effect, `ab`.

```{r}
post <-
  post %>% 
  mutate(ab = b_pmi_cond*b_reaction_pmi)
```

Now we've got `ab` as a properly computed vector, we can summarize it with the `quantile()` function.

```{r}
quantile(post$ab, probs = c(.5, .025, .975)) %>% 
  round(digits = 3)
```

And we can even visualize it as a density.

```{r, fig.width = 6, fig.height = 2}
post %>% 
  
  ggplot(aes(x = ab)) +
  geom_density(color = "transparent", 
               fill = colorblind_pal()(2)[2]) +
  scale_y_continuous(NULL, breaks = NULL) +
  labs(title = expression(paste("Our indirect effect, the ", italic("ab"), " pathway")),
       x = NULL) +
  theme_classic()
```





**More to come**






Note. The analyses in this document were done with:

* R           3.4.4
* RStudio     1.1.442
* rmarkdown   1.9
* readr       1.1.1
* tidyverse   1.2.1
* rstan       2.17.3
* brms        2.3.1
* ggthemes    3.5.0

## Reference

Hayes, A. F. (2018). *Introduction to mediation, moderation, and conditional process analysis: A regression-based approach.* (2nd ed.). New York, NY, US: The Guilford Press.